---
title: "TS2HW1 Code Doc"
author: "Blue Team 16"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
    df_print: paged
---

```{r, include=FALSE}
library(tidyverse)
library(rstudioapi)
library(tsibble)
library(fabletools)
library(scales)
library(fable)
library(forecast)
library(extrafont)
library(imputeTS)
library(feasts)
library(conflicted)
library(here)
setwd(here())
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
loadfonts(device = "win")
font_import()
```

# Setup
```{r}
# Reading in the data
energy <- read.csv("data/hrl_load_metered.csv")
energyTest <- read.csv("data/hrl_load_metered - test1.csv")

# Duplicate Spring removal
energy <- energy %>% group_by(datetime_beginning_ept) %>% 
  summarize(mw = mean(mw, na.rm = T),.groups = "drop") %>% 
  rename("time" = datetime_beginning_ept) 

energyTest <- energyTest %>% 
  group_by(datetime_beginning_ept) %>% 
  summarize(mw = mean(mw, na.rm = T),.groups = "drop") %>% 
  rename("time" = datetime_beginning_ept) 

# Fixing time values 
energy$time <- as.POSIXct(energy$time,format = "%m/%d/%y %H:%M",tz = "America/New_York")

energyTest$time <- as.POSIXct(energyTest$time,format = "%m/%d/%y %H:%M",tz = "America/New_York")

# Converting to Tsibble and filling gaps

energy <- energy %>% as_tsibble(index = time) %>% fill_gaps()

energyTest <- energyTest %>% as_tsibble(index = time) %>% fill_gaps()

# Imputing fall
energy <- energy %>% na_interpolation(option = "spline")

energyTest <- energyTest %>% na_interpolation(option = "spline")

## DUMMY VARIABLE MODEL HELPERS

# Add set indicator for separating
energy <- energy %>% mutate(set = "train")
energyTest <- energyTest %>% mutate(set = "test")

# Add all helper variables and combine sets to ensure factor levels are present throughout
energyCombined <- bind_rows(energy,energyTest) %>% 
  mutate(month = factor(month(time))) %>%
  mutate(hour = factor(hour(time))) %>% 
  mutate(dayOfWeek = factor(wday(time))) %>% 
  mutate(year = factor(year(time))) %>% 
  mutate(day = factor(day(time)))

# Separate sets back out
energy <- energyCombined %>% filter(set == "train")
energyTest <- energyCombined %>% filter(set == "test")
```
# Investigating Series

## Stationarity
```{r}
gg_tsdisplay(energy[1:15000,])
```
  
  Very doubtful of stationary, this is real data

```{r}
ndiffs(energy$mw)
```

All models perform better with differences, and statistical tests indicate we should be taking a first difference. All plots and models are on differenced data going forward.



## Selecting Lags
```{r}
ggAcf(difference(energy$mw),lag.max = 100)

ggPacf(difference(energy$mw),lag.max = 100)
```

PACF:Seeing exponential decrease in PACF aside from two spikes at lags 1, 2, and 24. Thinking p = 2, and P = 1 for a Basic Arima model.  

ACF: Large spikes throughout. Hard to tell with such complex seasonality. Would say q = 3 to account for the large spikes at 1,2,3, and Q = 7, because looking back any farther than a week feels like it wouldnt be sound. 

# Exponential Smoothing Models

```{r,eval = F}
# Create all Exponential Smoothing models 
etsFit <- energy %>% mutate(mw = difference(mw))
  model(
    `SES` = ETS(mw ~ error("A") + trend("N") + season("N")),
    `Linear` = ETS(mw ~ error("A") + trend("A") + season("N")),
    `Damped Linear` = ETS(mw ~ error("A") + trend("Ad") + season("N")),
    `Holt-Winters Additive` = ETS(mw ~ error("A") + trend("A") + season("A")),
    `Holt-Winters' Multiplicative` = ETS(mw ~ error("M") + trend("A") + season("M")),
    `Holt-Winters' Multiplicative Damped` = ETS(mw ~ error("M") + trend("Ad") + season("M"))
  )
  
```

```{r,eval = F}
# Saving the file so I dont have to run these again...
save(etsFit,file = "etsFit.RData")
```

# Deterministic models

## Seasonal ARIMA 
```{r,eval = F}
# Basic Seasonal Search 

# Built according to the lag chart investigation

seasonalArima <- energy %>% 
  model(
    search1 = ARIMA(mw ~pdq(d = 1) + PDQ(D = 0),stepwise = T),
    seasonalArima1 = ARIMA(mw ~pdq(p = 2,d = 1,q = 3) + PDQ(P = 1,D = 0,Q = 7)),
    seasonalArima2 = ARIMA(mw ~pdq(p = 2,d = 1,q = 3) + PDQ(P = 1,D = 0,Q = 5)),
    search2 = ARIMA(mw ~pdq(p = 2,d = 1,q = 3) + PDQ(D = 0),stepwise = F),
  )
# Saving the file so I dont have to run these again...
save(seasonalArima,file = "seasonalArima.RData")
```

## Dummy Variable Models

```{r,include = F}
# Making an older version of the test dataset in order to forecast with an 
# earlier version of dummy variable models
energyTest2 <- energyTest %>% select(-day) %>% rename("day" = dayOfWeek)
energy2 <- energy %>% select(-day) %>% rename("day" = dayOfWeek)
```

```{r,eval = F}
## Models  are split into chunks to prevent 
## loss of progress due to crashes / reboots, etc.

# Older version of dummyVariable model
print(Sys.time())
time <- Sys.time()
dummyArima <- energy2 %>% 
  model(
    dayHour = ARIMA(
      mw ~ 1 + day + hour + 
        pdq(d = 1) + PDQ(D = 0)
    )
  )

print("dummyArima Saved")
print(Sys.time() - time)
save(dummyArima,file = "dummyArima.RData")


print(Sys.time())
time <- Sys.time()
# Model 1
dummyArima1 <- energy %>% 
  model(
    HourMonthSearch = ARIMA(
      mw ~ 1 + hour + month + 
        pdq(d = 1) + PDQ(D = 0)
    )
  )

# 2.7 hours
print("dummyArima1 Saved")
print(Sys.time() - time)
save(dummyArima1,file = "dummyArima1.RData")


time <- Sys.time()
# Model 2
dummyArima2 <- energy %>% 
  model(
    HourMonthdayOfWeekSearch = ARIMA(
      mw ~ 1 + hour + month + dayOfWeek + 
        pdq(d = 1) + PDQ(D = 0)
    )
  )

# 5.5 hours
print("dummyArima2 Saved")
print(Sys.time() - time)
save(dummyArima2,file = "dummyArima2.RData")

# Lots more here were run, but alot just would output as NULL MODEL 
```

## Fourier Models
```{r,eval = F}
 time <- Sys.time()
# Throwing the kitchen sink at "complex seasonality" adjustments
fourierModels <- energy %>% 
  model(
    dayMonthWeekYear1 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 10) + 
              fourier(period = 24 * 30,K= 3) + 
              fourier(period = 24 * 7, K = 5)+
              fourier(period = 24 * 365, K = 3)
    ) + 
      dayMonthWeekYear2 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 12) + 
              fourier(period = 24 * 30,K= 5) + 
              fourier(period = 24 * 7, K = 7)+
              fourier(period = 24 * 365, K = 5)
    ) + 
      dayMonthWeekYear3 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 12) + 
              fourier(period = 24 * 30,K= 10) + 
              fourier(period = 24 * 7, K = 10)+
              fourier(period = 24 * 365, K = 10)
    ) + 
      dayMonthYear2 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 10) + 
              fourier(period = 24 * 30,K= 5) + 
              fourier(period = 24 * 365, K = 5)
    )+ 
      dayMonthYear3 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 10) + 
              fourier(period = 24 * 30,K= 8) + 
              fourier(period = 24 * 365, K = 8)
    ) + 
      dayMonthYear4 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 10) + 
              fourier(period = 24 * 30,K= 10) + 
              fourier(period = 24 * 365, K = 10)
    ) + 
      dayMonthYear1 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 12) + 
              fourier(period = 24 * 30,K= 3) + 
              fourier(period = 24 * 365, K = 3)
    ) + 
      dayMonthYear5 = 
      ARIMA(
        mw ~ pdq(d = 1) + PDQ(0,0,0) + 
              fourier(period = 24,K = 12) + 
              fourier(period = 24 * 30,K= 12) + 
              fourier(period = 24 * 365, K = 12)
    )
  )
```


# Model Eval
## Loading in Models

Models are loaded in from files after they are run to ensure that they are saved properly.
```{r}
load("etsFit.RData")

load("seasonalArima.RData")

load("dummyArima1.RData")
load("dummyArima2.RData")
load("dummyArima.RData")

load("fourierModels.RData")
```



## Model Glancing

```{r}
models <- bind_cols(fourierModels,dummyArima,dummyArima1,dummyArima2,etsFit,seasonalArima)
models %>% glance() %>% select(.model,"AICc") %>% arrange(AICc)
```

Seasonal ARIMA models built according to lag charts fit the data the best.  
Dummy variable models are the second best option, but are significantly worse.

## White Noise Eval
```{r}
# Only looking at the top 4 models, arranged by AICc

# SeasonalArima1
models %>% select(seasonalArima1) %>% gg_tsresiduals() + ggtitle("SeasonalArima1")
models %>% select(seasonalArima1) %>% residuals() %>% ggPacf() + ggtitle("SeasonalArima1 Pacf")

# SeasonalArima2
models %>% select(seasonalArima2) %>% gg_tsresiduals() + ggtitle("SeasonalArima2")
models %>% select(seasonalArima2) %>% residuals() %>% ggPacf() + ggtitle("SeasonalArima2 Pacf")


# Dummy variable dayOfWeek, Hour of Day
models %>% select(dayHour) %>% gg_tsresiduals() + ggtitle("dayHour")
models %>% select(dayHour) %>% residuals() %>% ggPacf() + ggtitle("dayHour Pacf")

# Dummy varaible, dayOfWeek, Hour of Day, and Month 
models %>% select(HourMonthdayOfWeekSearch) %>% gg_tsresiduals() + ggtitle("HourMonthdayOfWeek")
models %>% select(HourMonthdayOfWeekSearch) %>% residuals() %>% ggPacf() + ggtitle("HourMonthdayOfWeek Pacf")

# Exponential Smoothing Model
etsFit %>% select("Holt-Winters' Multiplicative") %>% gg_tsresiduals() + ggtitle("Exponential Smoothing")
etsFit %>% select("Holt-Winters' Multiplicative") %>% residuals() %>% ggPacf() + ggtitle("Exponential Smoothing Pacf")
```

## Model Forecasting 
```{r}
etsForecasts <- etsFit %>% fabletools::forecast(new_data = energyTest) 

seasonalForecasts <- seasonalArima %>% fabletools::forecast(h= nrow(energyTest))

dummyForecasts1 <- dummyArima1 %>% fabletools::forecast(new_data = energyTest)

dummyForecasts2 <- dummyArima2 %>% fabletools::forecast(new_data = energyTest)

# Old dummy variable models, old dataset
dummyForecasts <- dummyArima %>% fabletools::forecast(new_data = energyTest2)

# Combining the forecasts
forecasts <- bind_rows(dummyForecasts,dummyForecasts1,dummyForecasts2,etsForecasts,seasonalForecasts)
```

## MAPE and MAE
```{r}
accuracies <- fabletools::accuracy(forecasts,energyTest) 

accuracies <- accuracies[,c(".model","MAPE","MAE")] %>% arrange(MAPE)

print(accuracies)
```

## Forecast Graphs

### Seasonal ARIMA
```{r}
graphForecast <- forecasts %>% filter(.model == "seasonalArima2")
graphForecast <- graphForecast$.mean
graphDF <- data.frame(
  forecast = graphForecast,
  actual = energyTest$mw,
  time = energyTest$time
)
mape <- accuracies %>% filter(.model == "seasonalArima2") %>% select(MAPE) %>% round(2)

plot <- ggplot(data = graphDF, aes(x = time)) + 
geom_line(aes(y = forecast, color = "Seasonal ARIMA\nForecast"), linewidth = 1.2,alpha =.7) + 
geom_line(aes(y = actual,color = "Observed"),linetype = "twodash",linewidth = 1.5) + 
scale_color_manual(values = c("Observed" = "black", "Seasonal ARIMA\nForecast" = "orange")) + 
scale_y_continuous(labels = comma) + 
labs(colour = "",
     title = paste0("Forecasted MW Usage for ","Seasonal ARIMA Model",""),
     subtitle = paste0("09/13/2024- 09/19/2024\n","MAPE: ",mape,"\n"),
     y ="Hourly\nMW",
     x = "") + 
theme_classic() + 
theme(axis.title.y = element_text(angle = 0,vjust = .5),
      text = element_text(family = "Twentieth Century for Kenmore"))

print(plot)
```



### Holt-Winters
```{r}
graphForecast <- forecasts %>% filter(.model == "Holt-Winters' Multiplicative")
graphForecast <- graphForecast$.mean
graphDF <- data.frame(
  forecast = graphForecast,
  actual = energyTest$mw,
  time = energyTest$time
)
mape <- accuracies %>% filter(.model == "Holt-Winters' Multiplicative") %>% select(MAPE) %>% round(2)

plot <- ggplot(data = graphDF, aes(x = time)) + 
geom_line(aes(y = forecast, color = "Holt-Winters \nForecast"), linewidth = 1.2,alpha =.7) + 
geom_line(aes(y = actual,color = "Observed"),linetype = "twodash",linewidth = 1.5) + 
scale_color_manual(values = c("Observed" = "black", "Holt-Winters \nForecast" = "orange")) + 
scale_y_continuous(labels = comma) + 
labs(colour = "",
     title = paste0("Forecasted MW Usage for ","Exponential Smoothing Model",""),
     subtitle = paste0("09/13/2024- 09/19/2024\n","MAPE: ",mape,"\n"),
     y ="Hourly\nMW",
     x = "") + 
theme_classic() + 
theme(axis.title.y = element_text(angle = 0,vjust = .5))

print(plot)
```