---
title: "Seasonal Models"
author: "Ethan"
date: "2024-10-03"
output: pdf_document
---

```{r,include = F}
library(rpart)
library(tidyverse)
library(rpart.plot)
library(datasets)
library(arules)
library(arulesViz)
library(ggplot2)
library(dplyr)
library(rpart)
library(rpart.plot)
library(TH.data)
library(ISLR2)
library(lattice)
library(stats)
library(rattle)
library(RColorBrewer)
library(caret)
library(ROCR)
library(tidyverse)  
library(cluster)  
library(factoextra) 
library(gridExtra)
library(NbClust)
library(dendextend)
library(class)
library(ClustOfVar)
library(gmodels)
library(DescTools)
library(ROCit)
library(survival)
library(MASS)
library(kableExtra)
library(partykit)
library(dbscan)
library(AmesHousing)
library(rstudioapi)
library(tidyverse)
library(rstudioapi)
library(fable.prophet)
library(tsibble)
library(fabletools)
library(scales)
library(fable)
library(forecast)
library(imputeTS)
library(feasts)
library(here)
library(conflicted)
library(reticulate)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

# Slide Deck 1 

## Intro

### Data Reading
```{r}
file.dir = "https://raw.githubusercontent.com/sjsimmo2/TimeSeries/master/"
input.file1 = "usairlines.csv"
USAirlines = read.csv(paste(file.dir, input.file1,sep = ""))
USAirlines <- USAirlines %>%
 mutate(date = yearmonth(lubridate::make_date(Year, Month)))
USAirlines_ts <- as_tsibble(USAirlines, index = date)
```

### Splitting
```{r}
train <- USAirlines_ts %>%
 select(Passengers, date, Month) %>%
 filter_index(~ "2007-03")
test <- USAirlines_ts %>%
 select(Passengers, date, Month) %>%
 filter_index("2007-04" ~ .)
dcmp <- USAirlines_ts %>%
 model(stl = STL(Passengers))
components(dcmp) %>% autoplot()
```

### Exponential Smoothing
```{r}
model_HW <- train %>%
 model(
 ETS(Passengers ~ error("M") + trend("A") + season("M"))
 )
model_HW_for <- model_HW %>%
 fabletools::forecast(h = 12)
fabletools::accuracy(model_HW_for, test)
```

## Seasonal

### Nsdiffs

```{r}
train %>%
 features(Passengers, unitroot_nsdiffs)
```
### Unit Root Test
```{r}
train %>%
 mutate(Pass_diff = difference(Passengers, lag = 12)) %>%
 features(Pass_diff, unitroot_ndiffs)
```

### Dummy Variable Model

```{r}
season_lin <- lm(Passengers ~ factor(Month), data = train)
summary(season_lin)

model_SD_ARIMA <- train %>%
 model(ARIMA(Passengers ~ factor(Month) + PDQ(0,0,0)))
report(model_SD_ARIMA)
```

### Fourier
```{r}
model_F_ARIMA <- train %>%
 model(ARIMA(Passengers ~ fourier(K = 6) + PDQ(0,0,0))
)
report(model_F_ARIMA)
```

## Stochastic

### Seasonal Differencing 
```{r}
# Using 
train %>%
 gg_tsdisplay(difference(Passengers, 12), plot_type = 'partial', lag = 36)
```


# Slide Deck 3 Prophet Model

## Making it 

```{r,eval = F}
# Making additonal holidays
holidays <- data.frame(
 holiday = 'Sep11',
 ds = as.Date(c('2001-09-01', '2001-10-01', '2001-11-01',
 '2001-12-01', '2002-01-01', '2002-02-01',
 '2002-09-01')),
 # Guessing this stops prophet from looking into the days around the ones we specify
 lower_window = 0,
 upper_window = 0
)

# Making the dataframe to build the model on
prophet.data <- data.frame(ds = seq(as.Date('1990-01-01'), as.Date('2007-03-01'), by = 'm'), y = train$Passengers)

# Add the holidays / point interventions we made
Prof <- prophet::prophet(holidays = holidays,weekly.seasonality = TRUE,daily.seasonality = TRUE)

# Add holidays up until 1995
Prof <- add_country_holidays(Prof, "US")

# Add a monthly seasonality with a prespecified period and fourier order
Prof <- add_seasonality(Prof, name='monthly', period=30.5, fourier.order=6)

# Fit the model
Prof <- fit.prophet(Prof, prophet.data)
```
## Forecasting Prophet
```{r}

#
forecast.data <- make_future_dataframe(Prof, periods = 12, freq = 'month')


plot(Prof, predict(Prof, forecast.data))

pred <- predict(Prof, forecast.data)

pred$yhat


Prophet.error <- test$Passengers - tail(pred$yhat,12)
Prophet.MAE <- mean(abs(Prophet.error))
Prophet.MAPE <- mean(abs(Prophet.error)/abs(test$Passengers))*100

Prophet.MAPE
```

# Slide Deck 4 Neural Network


## Making it
```{r}
train$Sep11 <- rep(0, 207)
train$Sep11[141] <- 1

set.seed(12345)
model_nnet <- train %>%
  mutate(diff_Pass = difference(Passengers, 12)) %>%
  model(
    # 
    hand = NNETAR(diff_Pass ~ Sep11 + AR(p = 1, P = 3)),
  
    auto = NNETAR(diff_Pass ~ Sep11)  
  )

model_nnet %>%
  select(auto) %>%
  report()
```
Will not Difference for us. Must difference

## Residual Investigation

### Hand
```{r}
model_nnet %>%
  select(hand) %>%
  gg_tsresiduals()
```

### Auto
```{r}
model_nnet %>%
  select(auto) %>%
  gg_tsresiduals()
```


## Forecasting

Since we took a difference, we must transform the forecasted values to non-differenced

```{r}
## Forecasting 
test$Sep11 <- rep(0, 12)
```

### Forecasting takes so long holy shit
```{r,eval = F}
model_nnet_for <- fabletools::forecast(model_nnet, test)
save(model_nnet_for,file = "model_nnet_for.RData")
```
### Adjusting differenced forecasts
```{r}
load("model_nnet_for.RData")

nnet_for_hand <- rep(NA, 12)
nnet_for_auto <- rep(NA, 12)

for(i in 1:12){
  nnet_for_hand[i] <- train$Passengers[length(train$Passengers) - 12 + i] + model_nnet_for$.mean[i]
}

for(i in 1:12){
  nnet_for_auto[i] <- train$Passengers[length(train$Passengers) - 12 + i] + model_nnet_for$.mean[i+12]
}
```

### Plotting forecasts

```{r}
Passenger <- ts(USAirlines$Passengers, start = 1990, frequency = 12)
plot_pass <- subset(Passenger, end = length(Passenger) - 12)

plot(plot_pass, main = "US Airline Passengers Forecasts", xlab = "Date", ylab = "Passengers (Thousands)", ylim = c(30000,80000), xlim = c(1990, 2009))
lines(ts(nnet_for_auto, start = c(2007, 4), frequency = 12), col = "blue")
lines(ts(nnet_for_hand, start = c(2007, 4), frequency = 12), col = "orange")
abline(v = 2007.25, col = "red", lty = "dashed")
```
Looks pretty amazing

### Calculating Mape and MAE

```{r}
NN_error_auto <- test$Passengers - nnet_for_auto

NN_MAE_auto <- mean(abs(NN_error_auto))
NN_MAPE_auto <- mean(abs(NN_error_auto)/abs(test$Passengers))*100

NN_error_hand <- test$Passengers - nnet_for_hand

NN_MAE_hand <- mean(abs(NN_error_hand))
NN_MAPE_hand <- mean(abs(NN_error_hand)/abs(test$Passengers))*100

NN_MAE_auto
```

```{r}
NN_MAPE_auto
```
Oh my god its insnae

# Slide Deck 5 - Ensemble Models

```{r}
Pass.Fit.NN <- rep(NA, 207)
NN.Model = model_nnet %>% select(auto)
fitted(NN.Model)
for(i in 25:207){
 Pass.Fit.NN[i] <- train[i - 24] + NN.Model$fitted[i - 12]
}
Pass.Fit.ARIMA <- Full.ARIMA$fitted
Pass.Fit.HWES <- HWES.USAir.train$fitted
Pass.Fit.Prophet <- head(predict(Prof, forecast.data)$yhat, 207)

## This doesnt work exactly, but this code restricts all the coefficients to sum to one.
## You set up your predictors as the forecasts, and then use the weights to average
## Seems stupid, is stupid, I dont like it.
WC.Model <- lm(training ~ offset(Pass.Fit.HWES) +
 I(Pass.Fit.ARIMA - Pass.Fit.HWES) +
 I(Pass.Fit.NN - Pass.Fit.HWES) +
 I(Pass.Fit.Prophet - Pass.Fit.HWES) - 1)
summary(WC.Model)

```
